---
title: "Neighborhood analysis"
author: "Katie Clary"
date: "3/2/2018"
output:
  html_document:
    code_folding: hide
    highlight: tango
    theme: cerulean
  word_document: default
---
```{r setup, include = F}
knitr::opts_knit$set(root.dir = "/Volumes/T1000/Analysis/kforthman/Neighborhood_Analysis/Tract")
```

```{r, message = F, include = F}
library(acs)
library(stringr)
library(e1071)
library(psych)
library(DMwR)
library(randomcoloR)
library(corrplot)
library(ggplot2)
library(lavaan)
library(scales)
library(RColorBrewer)
library(GPArotation)
```

```{r, include = F}
factorName_n <- function(factor){
  c("African\nAmericans\nin Tract", 
    "Seniors\nin Tract", 
    "Singletons\nin Tract", 
    "Affluence", 
    "Noncitizens\nin Tract")[as.numeric(substr(factor, 3,3))]
}
factorName <- function(factor){
  c("African Americans in Tract", 
    "Seniors in Tract", 
    "Singletons in Tract", 
    "Affluence", 
    "Noncitizens in Tract")[as.numeric(substr(factor, 3,3))]
}
stateName <- read.csv("data/state_code.csv")
stateCodes <- str_pad(as.character(stateName[,2]), 2, "left", pad = " ")
```

```{r, collapse=F}
endyear = 2015
nfactors = 5
```

# Selecting variables
Alters the original matrix to be more suited for the factor analysis. Renames variables, throws out certain variables, and alters some variables to be percentages. Additionaly, gets rid of all tracts with a population of zero.
```{r, eval = F, include=T}

# Alters the original matrix to be more suited for the factor analysis.
# Renames variables, throws out certain variables, and alters some variables to be percentages.
# Additionaly, gets rid of all tracts with a population of zero.

# Read in the original matrix.
original <- read.csv(paste0("output/2015/useful_ACS_originalData.csv"), row.names = 1)

# If the state code is only a single digit, it is padded.
rownames(original) <- str_pad(rownames(original), 11, "left", pad="0")

# Define the number of tracts
n <- dim(original)[1]

# --- Now, begin adding variables to a new matrix and changing them as needed. --- #

# Total population ----
condensed <- original['Total']

# Age and gender ----
Males.age <- c("Male...Under.5.years","Male...5.to.9.years","Male...10.to.14.years","Male...15.to.17.years",
               "Male...18.and.19.years","Male...20.years","Male...21.years","Male...22.to.24.years",
               "Male...25.to.29.years","Male...30.to.34.years","Male...35.to.39.years","Male...40.to.44.years",
               "Male...45.to.49.years","Male...50.to.54.years","Male...55.to.59.years","Male...60.and.61.years",
               "Male...62.to.64.years","Male...65.and.66.years","Male...67.to.69.years","Male...70.to.74.years",
               "Male...75.to.79.years","Male...80.to.84.years","Male...85.years.and.over")

Females.age <- c("Female...Under.5.years","Female...5.to.9.years","Female...10.to.14.years",
                 "Female...15.to.17.years","Female...18.and.19.years","Female...20.years",
                 "Female...21.years","Female...22.to.24.years","Female...25.to.29.years",
                 "Female...30.to.34.years","Female...35.to.39.years","Female...40.to.44.years",
                 "Female...45.to.49.years","Female...50.to.54.years","Female...55.to.59.years",
                 "Female...60.and.61.years","Female...62.to.64.years","Female...65.and.66.years",
                 "Female...67.to.69.years","Female...70.to.74.years","Female...75.to.79.years",
                 "Female...80.to.84.years","Female...85.years.and.over")

Age <- original[Males.age]+original[Females.age]
for(i in 1:n){
  Age[i,] <- c(2.5,7,12,16,18.5,20,21,23,27,32,37,42,47,52,57,60.5,63,65.5,68,72,77,82,85) * Age[i,]
}
Age <- rowSums(Age)
Age <- Age/original["Total..Sex.by.age"]
condensed['Age'] <- Age

# Race ----
condensed['Total..Race'] <- original['Total...Race']

condensed['White'] <- original['White.alone']/original['Total...Race']

condensed['Black'] <- original['Black.or.African.American.alone']/original['Total...Race']

condensed['Asian'] <- original['Asian.alone']/original['Total...Race']

condensed['Some.other.race'] <- ((original['Some.other.race.alone'] + 
                                    original['Native.Hawaiian.and.Other.Pacific.Islander.alone'] +
                                    original['American.Indian.and.Alaska.Native.alone'])
                                 /original['Total...Race'])

# Ethnicity ----
condensed['Total..Hispanic.or.Latino'] <- original['Total...Hispanic.or.Latino.by.Specific.Origin']

condensed['Hispanic.or.Latino'] <- original['Hispanic.or.Latino.']/original['Total...Hispanic.or.Latino.by.Specific.Origin']

# Nativity, citizenship, and mobility ----
condensed['Total...Nativity.and.Citizenship.Status'] <- original['Total...Nativity.and.Citizenship.Status.in.the.United.States']

condensed['Not.US.citizen'] <- ((original['Not.a.U.S..citizen'] 
                                 /original['Total...Nativity.and.Citizenship.Status.in.the.United.States']))

# Mobility ----

condensed['Total...Geographical.Mobility.in.the.Past.Year'] <- original['Total...Geographical.Mobility.in.the.Past.Year.by.Citizenship.Status.for.Current.Residence.in.the.U.S.']

condensed['Moved.in.past.year'] <- 1 - (original['Same.house.1.year.ago.']/original['Total...Geographical.Mobility.in.the.Past.Year.by.Citizenship.Status.for.Current.Residence.in.the.U.S.'])

# Transportation to work ----
condensed['Total..Means.of.Transportation.to.Work'] <- original['Total...Means.of.Transportation.to.Work']

condensed['Drove'] <- ((original['Car..truck..or.van.']+
                          original['Motorcycle']) / 
                         original['Total...Means.of.Transportation.to.Work'])

condensed['Bicycle.or.walked'] <- ((original['Bicycle'] + original['Walked']) / 
                                     original['Total...Means.of.Transportation.to.Work'])

# Time left for work ----
condensed['Total..Time.Leaving.Home.to.Go.to.Work'] <- original['Total...Time.Leaving.Home.to.Go.to.Work']

Time.Leaving.Home.to.Go.to.Work <- c("X12.00.a.m..to.4.59.a.m.", "X5.00.a.m..to.5.29.a.m.", "X5.30.a.m..to.5.59.a.m.", "X6.00.a.m..to.6.29.a.m.", 
                                     "X6.30.a.m..to.6.59.a.m.", "X7.00.a.m..to.7.29.a.m.", "X7.30.a.m..to.7.59.a.m.", "X8.00.a.m..to.8.29.a.m.", 
                                     "X8.30.a.m..to.8.59.a.m.", "X9.00.a.m..to.9.59.a.m.", "X10.00.a.m..to.10.59.a.m.", "X11.00.a.m..to.11.59.a.m.", 
                                     "X12.00.p.m..to.3.59.p.m.", "X4.00.p.m..to.11.59.p.m.")
Time.Leaving.Home.to.Go.to.Work <- original[Time.Leaving.Home.to.Go.to.Work]
for(i in 1:n){
  Time.Leaving.Home.to.Go.to.Work[i,] <- c(0,5,5.5,6,6.5,7,7.5,8,8.5,9,10,11,12,16) * Time.Leaving.Home.to.Go.to.Work[i,]
}
Time.Leaving.Home.to.Go.to.Work <- rowSums(Time.Leaving.Home.to.Go.to.Work)
Time.Leaving.Home.to.Go.to.Work <- Time.Leaving.Home.to.Go.to.Work/original['Total...Time.Leaving.Home.to.Go.to.Work']
condensed['Time.leaving.home'] <- Time.Leaving.Home.to.Go.to.Work

# Time spent traveling to work ----
condensed['Total..Travel.Time.to.Work'] <- original['Total...Travel.Time.to.Work']

Travel.Time.to.Work <- c("Less.than.5.minutes", "X5.to.9.minutes", "X10.to.14.minutes", 
                         "X15.to.19.minutes", "X20.to.24.minutes", "X25.to.29.minutes", 
                         "X30.to.34.minutes", "X35.to.39.minutes", "X40.to.44.minutes",
                         "X45.to.59.minutes", "X60.to.89.minutes", "X90.or.more.minutes")
Travel.Time.to.Work <- original[Travel.Time.to.Work]
for(i in 1:n){
  Travel.Time.to.Work[i,] <- c(3, 7, 12, 17, 22, 27, 32, 37, 42, 47, 74.5, 90) * Travel.Time.to.Work[i,]
}
Travel.Time.to.Work <- rowSums(Travel.Time.to.Work)
Travel.Time.to.Work <- Travel.Time.to.Work/original['Total...Travel.Time.to.Work']
condensed['Travel.time.to.work'] <- Travel.Time.to.Work

# Household type ----
condensed['Total..Household.Type'] <- original['Total...Household.Type..including.Living.Alone.']

condensed['Single.moms'] <- original['Family.households...Other.family...Female.householder..no.husband.present']/original['Total...Household.Type..including.Living.Alone.']

condensed['Lives.alone'] <- original['Nonfamily.households...Householder.living.alone']/original['Total...Household.Type..including.Living.Alone.']

condensed['Nonfamily.household'] <- original['Nonfamily.households...Householder.not.living.alone']/original['Total...Household.Type..including.Living.Alone.']

# Marital status ----
condensed['Total..Marital.Status.for.the.Population.15.Years.and.over'] <- original['Total...Sex.by.Marital.Status.for.the.Population.15.Years.and.over']

Males.marital.status <- c("Male...Never.married", 
                          "Male...Now.married...Married..spouse.present",
                          "Male...Now.married...Married..spouse.absent.", 
                          "Male...Widowed", "Male...Divorced")

Females.marital.status <- c("Female...Never.married", 
                            "Female...Now.married...Married..spouse.present",
                            "Female...Now.married...Married..spouse.absent.", 
                            "Female...Widowed", "Female...Divorced")

Males.marital.status <- original[Males.marital.status]
Females.marital.status <- original[Females.marital.status]

Marital.status <- Males.marital.status + Females.marital.status

names(Marital.status) <- c("Never.married", 
                           "Married.spouse.present", 
                           "Married.spouse.absent", 
                           "Widowed", "Divorced")

for(i in 1:ncol(Marital.status)){
  Marital.status[,i] <- Marital.status[,i]/original['Total...Sex.by.Marital.Status.for.the.Population.15.Years.and.over']
}

condensed[names(Marital.status)] <- Marital.status



# Education Level ----
condensed['Total...Educational.Attainment.for.the.Population.25.Years.and.Over'] <- original['Total...Educational.Attainment.for.the.Population.25.Years.and.Over']

Educational.Attainment <- c("No.schooling.completed", "Nursery.school", "Kindergarten", "X1st.grade", "X2nd.grade",
                            "X3rd.grade", "X4th.grade", "X5th.grade", "X6th.grade", "X7th.grade",
                            "X8th.grade", "X9th.grade", "X10th.grade", "X11th.grade", "X12th.grade..no.diploma",
                            "GED.or.alternative.credential", "Regular.high.school.diploma", "Some.college..less.than.1.year",
                            "Some.college..1.or.more.years..no.degree", "Associate.s.degree",
                            "Bachelor.s.degree", "Master.s.degree", "Professional.school.degree", "Doctorate.degree")

Educational.Attainment <- original[Educational.Attainment]

for(i in 1:n){
  Educational.Attainment[i,] <- ((0 * sum(Educational.Attainment[i,1:15]) +
                                    1 * sum(Educational.Attainment[i,16:20]) +
                                    2 * sum(Educational.Attainment[i,21:24]))
                                 /original[i, 'Total...Educational.Attainment.for.the.Population.25.Years.and.Over'])
}

Educational.Attainment <- Educational.Attainment[,2]

condensed['Education'] <- Educational.Attainment

# Disability status ----
condensed['Total...Sex.by.Age.by.Disability.Status'] <- original['Total...Sex.by.Age.by.Disability.Status']

Male.withDisability <- c("Male...Under.5.years...With.a.disability",
                         "Male...5.to.17.years...With.a.disability",
                         "Male...18.to.34.years...With.a.disability",
                         "Male...35.to.64.years...With.a.disability",
                         "Male...65.to.74.years...With.a.disability",
                         "Male...75.years.and.over...With.a.disability")

Female.withDisability <- c("Female...Under.5.years...With.a.disability",
                           "Female...5.to.17.years...With.a.disability",
                           "Female...18.to.34.years...With.a.disability",
                           "Female...35.to.64.years...With.a.disability",
                           "Female...65.to.74.years...With.a.disability",
                           "Female...75.years.and.over...With.a.disability")

Male.withDisability <- original[Male.withDisability]
Female.withDisability <- original[Female.withDisability]

Male.withDisability <- rowSums(Male.withDisability)
Female.withDisability <- rowSums(Female.withDisability)

Disabled <- (Male.withDisability + Female.withDisability)/original['Total...Sex.by.Age.by.Disability.Status']

condensed['Disabled'] <- Disabled

# Income ----
condensed['Total..Household.Income.in.the.Past.12.Months..in.2015.Inflation.Adjusted.Dollars.'] <- original['Total..Household.Income.in.the.Past.12.Months..in.2015.Inflation.Adjusted.Dollars.']

Household.Income <- c("Less.than..10.000", "X.10.000.to..14.999", "X.15.000.to..19.999", 
                      "X.20.000.to..24.999", "X.25.000.to..29.999", "X.30.000.to..34.999", 
                      "X.35.000.to..39.999", "X.40.000.to..44.999", "X.45.000.to..49.999",
                      "X.50.000.to..59.999", "X.60.000.to..74.999", "X.75.000.to..99.999", 
                      "X.100.000.to..124.999", "X.125.000.to..149.999", "X.150.000.to..199.999", 
                      "X.200.000.or.more")

Household.Income <- original[Household.Income]

for(i in 1:n){
  Household.Income[i,] <- c(4999.5,12499.5,17499.5,22499.5,27499.5,32499.5,37499.5,42499.5,
                            47499.5,54999.5,67499.5,87499.5,112499.5,137499.5,174999.5,200000) * Household.Income[i,]
}
Household.Income <- rowSums(Household.Income)
Household.Income <- Household.Income/original['Total..Household.Income.in.the.Past.12.Months..in.2015.Inflation.Adjusted.Dollars.']
condensed['Income'] <- Household.Income

# Social security ----
condensed['Total..Social.Security.Income.for.Households'] <- original['Total...Social.Security.Income.for.Households']

condensed['w.Social.Security'] <- original['With.Social.Security.income']/original['Total...Social.Security.Income.for.Households']

# Supplementary security income ----
condensed['Total..Supplemental.Security.Income.for.Households'] <- original['Total...Supplemental.Security.Income..SSI..for.Households']

condensed['w.SSI'] <- original['With.Supplemental.Security.Income..SSI.']/original['Total...Supplemental.Security.Income..SSI..for.Households']

# Public assistance ----
condensed['Total..Public.Assistance.Income.for.Households'] <- original['Total...Public.Assistance.Income.for.Households']

condensed['w.PAI'] <- original['With.public.assistance.income']/original['Total...Public.Assistance.Income.for.Households']

# Retirement income ----
condensed['Total..Retirement.Income.for.Households'] <- original['Total...Retirement.Income.for.Households']

condensed['w.Retirement.income'] <- original['With.retirement.income']/original['Total...Retirement.Income.for.Households']

condensed['Retirement.income'] <- original['Per.capita.income.in.the.past.12.months..in.2015.Inflation.adjusted.dollars.']

# Employment status ----
condensed['Total...Employment.Status.for.the.Population.16.Years.and.Over'] <- original['Total...Employment.Status.for.the.Population.16.Years.and.Over']

condensed['Unemployed'] <- original['In.labor.force...Civilian.labor.force...Unemployed']/original['Total...Employment.Status.for.the.Population.16.Years.and.Over']

condensed['Not.in.labor.force'] <- original['Not.in.labor.force']/original['Total...Employment.Status.for.the.Population.16.Years.and.Over']
# Occupancy status ----
condensed['Total..Occupancy.Status'] <- original['Total...Occupancy.Status']

condensed['Vacant'] <- original['Vacant']/original['Total...Occupancy.Status']

# Tenure ----
condensed['Total..Tenure'] <- original['Total...Tenure']

condensed['Not.owner.occupied'] <- 1 - (original['Owner.occupied']/original['Total...Tenure'])

# Number of rooms ----
condensed['Median.number.of.rooms'] <- original['Median.number.of.rooms']

# Units in structure, home type, and home age ----
condensed['Total..Units.in.Structure'] <- original['Total...Units.in.Structure']

Units.in.structure <- c("X1..detached", "X1..attached", "X2", "X3.or.4", "X5.to.9", "X10.to.19", "X20.to.49", "X50.or.more")
Units.in.structure <- original[Units.in.structure]
for(i in 1:n){
  Units.in.structure[i,] <- c(1, 1, 2, 3.5, 7, 14.5, 34.5, 50) * Units.in.structure[i,]
}
Units.in.structure <- rowSums(Units.in.structure)
condensed['Units.in.structure'] <- Units.in.structure / (original['Total...Units.in.Structure']-
                                                           original['Mobile.home'] - original['Boat..RV..van..etc.'])


condensed['Mobile.home'] <- (original['Mobile.home'] + original['Boat..RV..van..etc.']) / original['Total...Units.in.Structure']

# Kitchen facilities ----
condensed['Total..Kitchen.facilities'] <- original['Total...KITCHEN.FACILITIES.FOR.ALL.HOUSING.UNITS']

condensed['Incomplete.kitchen'] <- 1 - (original['Complete.kitchen.facilities']/original['Total...KITCHEN.FACILITIES.FOR.ALL.HOUSING.UNITS'])

# Rent cost ----
condensed['Total..Contract.Rent'] <- original['Total..Contract.Rent']

Rent <- c("With.cash.rent...Less.than..100", "With.cash.rent....100.to..149", "With.cash.rent....150.to..199", 
          "With.cash.rent....200.to..249", "With.cash.rent....250.to..299", "With.cash.rent....300.to..349",
          "With.cash.rent....350.to..399", "With.cash.rent....400.to..449", "With.cash.rent....450.to..499", 
          "With.cash.rent....500.to..549", "With.cash.rent....550.to..599", "With.cash.rent....600.to..649",
          "With.cash.rent....650.to..699", "With.cash.rent....700.to..749", "With.cash.rent....750.to..799", 
          "With.cash.rent....800.to..899", "With.cash.rent....900.to..999", "With.cash.rent....1.000.to..1.249",
          "With.cash.rent....1.250.to..1.499", "With.cash.rent....1.500.to..1.999", "With.cash.rent....2.000.to..2.499", 
          "With.cash.rent....2.500.to..2.999", "With.cash.rent....3.000.to..3.499", "With.cash.rent....3.500.or.more")
Rent <- original[Rent]
for(i in 1:n){
  Rent[i,] <- c(49.5,124.5,174.5,224.5,274.5,324.5,374.5,424.5,474.5,524.5,574.5,624.5,674.5,724.5,774.5,849.5,949.5,1124.5,
                1374.5,1749.5,2249.5,2749.5,3249.5,3500) * Rent[i,]
}
Rent <- rowSums(Rent)
Rent <- Rent/original['With.cash.rent.']
condensed['Cash.rent.cost'] <- Rent

# Home value ----
condensed['Total..Value'] <- original['Total..Value']

Home.value <- c("Less.than..10.000.1", "X.10.000.to..14.999.1", "X.15.000.to..19.999.1", "X.20.000.to..24.999.1", "X.25.000.to..29.999.1", "X.30.000.to..34.999.1", "X.35.000.to..39.999.1",
                "X.40.000.to..49.999", "X.50.000.to..59.999.1", "X.60.000.to..69.999", "X.70.000.to..79.999", "X.80.000.to..89.999", "X.90.000.to..99.999", "X.100.000.to..124.999.1",
                "X.125.000.to..149.999.1", "X.150.000.to..174.999", "X.175.000.to..199.999", "X.200.000.to..249.999", "X.250.000.to..299.999", "X.300.000.to..399.999", "X.400.000.to..499.999",
                "X.500.000.to..749.999", "X.750.000.to..999.999", "X.1.000.000.to..1.499.999", "X.1.500.000.to..1.999.999", "X.2.000.000.or.more")
Home.value <- original[Home.value]
for(i in 1:n){
  Home.value[i,] <- c(4999.5,12499.5,17499.5,22499.5,27499.5,32499.5,37499.5,44999.5,54999.5,64999.5,
                      74999.5,84999.5,94999.5,112499.5,137499.5,162499.5,187499.5,224999.5,274999.5,
                      349999.5,449999.5,624999.5,874999.5,1250000,1750000,2000000) * Home.value[i,]
}
Home.value <- rowSums(Home.value)
Home.value <- Home.value/original['Total..Value']
condensed['Home.value'] <- Home.value

# Morgage status ----
condensed['Total..Mortgage.status'] <- original['Total...MORTGAGE.STATUS']

condensed['Mortgage'] <- original['Housing.units.with.a.mortgage..contract.to.purchase..or.similar.debt.']/original['Total...MORTGAGE.STATUS']

# ----

# Remove all tracts with a total population of zero.
condensed <- condensed[which(condensed[,"Total"] > 0),]
# Get rid of vectors containing Totals
condensed <- condensed[,!grepl("Total", names(condensed))]

# Create the cleaned up matrix
#write.csv(condensed, paste0("useful_ACS_condensedData.csv"))

```

# Transforming variables
This chunk finds the optimal transformation for normalization of each of the variables and outputs a matrix of the transformed data.
```{r, eval = F, include=T}

# This function finds the optimal transformation for normalization of each of the variables
# and outputs a matrix of the transformed data.

# Grab the altered ACS matrix
mydata <- read.csv("useful_ACS_condensedData.csv", row.names = 1)

# Define the number of tracts and the number of variables.
n <- dim(mydata)[1]
nvar <- dim(mydata)[2]

# Create an empty matrix to store the transformed variables.
trans_data <- matrix(nrow = n, ncol = 0)

# n_trans_val is the number of different strengths of transformation we want to test
# for getting the most normal curve.
n_trans_val <- 50

# Data is 'bent' so that is is curved in a particular direction.
# Right skewed data will be bent so that they curve concave down.
# Left skewed data will be bent so that they curve concave up.
for(i in 1:nvar){
  var_obs <- mydata[,i]
  
  var_obs_skewness <- round(skewness(var_obs, na.rm = T), 1)
  
  m <- max(var_obs, na.rm = T)
  # The transformation values alter the strength of the transformation. The larger the 
  # trans_val is, the less strong the transformation will be.
  trans_val <- seq(0, m, length.out = n_trans_val+1)
  trans_val <- trans_val[2:(n_trans_val+1)]
  # The trans values are given a logarithmic step scale. This is because
  # smaller values have a larger incremental effect on the transformation
  # strength.
  trans_val <- (m*(exp(5*trans_val/m) - 1)) / (exp(5) - 1)
  
  # Now, the function does one transformation with variables skewed to the right
  # and the opposite transformation for variables skewed to the left.
  if(var_obs_skewness > 1){
    
    # The function that transforms the variables has different effects for
    # different values of the trans_val. For smaller values of trans_val,
    # the transformation function will have a stronger effect on the variable's
    # distribution.
    trans_func <- function(trans_val){
      h <- m/(log(m+trans_val)-log(trans_val))
      var_obs_trans <- h*(log(var_obs + trans_val) - log(trans_val))
      return(var_obs_trans)
    }
    
    # Each column of the following matrix is a transformation of a different strength.
    var_obs_trans_mat <- sapply(trans_val, trans_func)
    colnames(var_obs_trans_mat) <- trans_val #round(trans_val, 1)
    # The kurtosis is found for each strength of transformation.
    kurt <- sapply(seq(1,n_trans_val), function(x){kurtosis(var_obs_trans_mat[,x], na.rm = T)})
    
    kurt_min_loc <- which.min(abs(kurt))
    var_obs_trans <- as.matrix(var_obs_trans_mat[,kurt_min_loc])
    
  }else if(var_obs_skewness < -1){
    
    # This function is the same as the one above, but flipped on the y=x axis.
    trans_func <- function(trans_val){
      h <- m/(log(m+trans_val)-log(trans_val))
      var_obs_trans <-  exp(var_obs/h + log(trans_val)) - trans_val
      return(var_obs_trans)
    }
    
    var_obs_trans_mat <- sapply(trans_val, trans_func)
    colnames(var_obs_trans_mat) <- trans_val #round(trans_val, 1)
    kurt <- sapply(seq(1,n_trans_val), function(x){kurtosis(var_obs_trans_mat[,x], na.rm = T)})
    
    kurt_min_loc <- which.min(abs(kurt))
    var_obs_trans <- as.matrix(var_obs_trans_mat[,kurt_min_loc])
    
  }else{
    
    # If the variable is not that skewed, it is not transformed.
    var_obs_trans <- var_obs
  }
  
  # If the observations are supposed to be of zero mean and unit variance, they are scaled.
  var_obs_trans <- scale(var_obs_trans)
  
  # Binds the transformed variable to a new matrix.
  trans_data <- cbind(trans_data, matrix(var_obs_trans))
}

trans_data <- knnImputation(trans_data)

colnames(trans_data) <- names(mydata)
rownames(trans_data) <- rownames(mydata)
#write.csv(trans_data, "useful_ACS_transformedData.csv")
```

# Factor Analysis
This chunk does the exploratory factor analysis.
```{r}
# ------- Import dataset ------- #
useful_ACS_transformedData <- as.matrix(read.csv("useful_ACS_transformedData.csv", row.names = 1))

# ------- Maximum Likelihood Factor Analysis ------- #
# entering data and extracting factors, 
# with varimax rotation 

# Create fit using PSYCH package
FA_Results <- fa(useful_ACS_transformedData, nfactors = nfactors, rotate="oblimin", fm = 'minres')
print(FA_Results, digits=2, cutoff=.3, sort=TRUE)
```

# Bootstrap resampling
```{r, eval = F}
p <- dim(FA_Results$loadings)[1]

colnames(FA_Results$loadings) <- factorName(colnames(FA_Results$loadings))

list_bootload <- list()

for(i in 1:2000){
  my_samp <- sample.int(nrow(useful_ACS_transformedData), nrow(useful_ACS_transformedData), replace = TRUE)
  
  my_boot <- useful_ACS_transformedData[my_samp,]
  
  #write.csv(my_boot, paste0("bootstrap/data/bootstrap_", i, ".csv"))
  
  print(paste("Matrix", i, "saved."))
  
  boot_fa <- fa(my_boot, nfactors = nfactors, rotate= "oblimin", fm = 'minres')
  
  print(paste("FA", i, "complete."))
  
  all_boot_cor <- matrix(nrow = nfactors, ncol = nfactors)
  
  for(j in 1:nfactors){
    for(k in 1:nfactors){
      all_boot_cor[j,k] <- cor(FA_Results$loadings[,j], boot_fa$loadings[,k])
    }
  }
  
  my.order <- sapply(c(1:nfactors), function(x){which.max(all_boot_cor[,x])})
  
  colnames(boot_fa$loadings) <- colnames(FA_Results$loadings)[my.order]
  
  #saveRDS(boot_fa, paste0("bootstrap/fa/bootstrap_", i, ".rds"))
  
  list_bootload[[i]]  <- boot_fa$loadings[,c("Racial Homogeneity", "Senoirity", "Singleton", "Affluence", "Citizenship")]
}

sum <- matrix(0, nrow = p, ncol = nfactors)
for (i in 1:2000){
  sum <- sum + list_bootload[[i]]
}

mean <- sum/2000

sum2 <- matrix(0, nrow = p, ncol = nfactors)
for (i in 1:2000){
  sum2 <- sum2 + (list_bootload[[i]] - mean)^2
}

se <- sqrt(sum2/1999)

#write.csv(se, "bootstrap_error.csv")
```

# Plotting the Factor Structure
Plots the factor structure for the given model.
```{r, fig.height=10, fig.width=10}
error <- read.csv("bootstrap_error.csv", row.names = 1)

# Each element in the following list stores a plot for one of the factors.
myplots <- list()
for(i in 1:nfactors){
  p <- length(FA_Results$loadings[,i])
  
  mylabel <- names(FA_Results$loadings[,i])
  labx <- seq(1, p)
  laby <- seq(1.1, 1.1, length.out = p)
  bgy <- seq(1, 1, length.out = p)
  laba <- seq((90 - 360/p/2), by = -360/p, length.out = p) + c(seq(0, 0, length.out = floor(p/2)), seq(180, 180, length.out = ceiling(p/2)))
  labj <-c(seq(0, 0, length.out = floor(p/2)), seq(1, 1, length.out = ceiling(p/2)))
  
  spoke_length <- c(1,5,1,1,4,3,5,1,1,6,2,6,3)
  
  sll <- length(spoke_length)
  mycolors <- grDevices::rainbow(sll)
  mycolors <- mycolors[c(1,8,2,9,3,10,4,11,5,12,6,13,7)]
  spoke_x <- c(0.5)
  for(j in 2:sll){
    spoke_x[j] <- spoke_x[j-1] + spoke_length[j-1]
  }
  
  varAlpha <- ifelse(abs(FA_Results$loadings[,i]) > 0.3, 1, 0.4)
  varCol <- ifelse(FA_Results$loadings[,i] > 0, "turquoise1", "violetred1")
  bgCol <- rep(c("grey92", "grey94", "grey96"), p)[1:p]
  this_error <- error[,i]
  
  print(factorName_n(colnames(FA_Results$scores)[i]))
  print(FA_Results$loadings[,i])
  
  df <- data.frame(var = seq(1,p), load = abs(FA_Results$loadings[,i]) , se = this_error)
  
  text.size <- 5
  
  ggp <- ggplot(df) + 
    geom_bar(aes(x = var, y = bgy), stat = "identity", width = 1, fill = bgCol) + 
    geom_bar(aes(x = var, y = load), stat = "identity", width = 1, fill = varCol, 
             alpha = varAlpha) + 
    geom_errorbar(aes(x = var, ymin = load - se, ymax = load + se), width = 0.5, color = "grey50", 
                  size = text.size/12) +
    scale_y_continuous(limits = c(-0.4, 1.5)) +
    annotate("text", labx, laby, label = mylabel, angle = laba, 
             size = text.size, col = "black", hjust = labj, alpha = varAlpha) +
    annotate("text", 1, -0.4, 
             label = paste0(factorName_n(colnames(FA_Results$scores)[i]), 
                            "\n(", as.character(100*round(FA_Results$Vaccounted[2,i],2)), 
                            "%)"),
             size = text.size*4/3, col = "black") +
    annotate("segment", x = spoke_x, xend = spoke_x + spoke_length, y = 1.025, yend = 1.025, 
             color = mycolors, size = text.size/4) +
    theme(plot.margin = margin(t = 0, r = 0, b = 0, l = 0, unit = "pt"), 
          axis.ticks=element_blank(),
          axis.title=element_blank(),
          axis.text=element_blank(), 
          panel.grid = element_blank(),
          panel.background = element_rect(fill = "white")) +
    coord_polar()
  
  plot(ggp)
  
}

```

# Comparing Mental Health to Neighborhood Factors
```{r}
cdc_data <- read.csv("500_Cities__Census_Tract-level_Data__GIS_Friendly_Format_.csv")
```
```{r, fig.height=10, fig.width=10}

t1 <- as.data.frame(cdc_data[,c("TractFIPS", "MHLTH_CrudePrev")])

factor_scores <- cbind(rownames(FA_Results$scores), FA_Results$scores)
t2 <- as.data.frame(factor_scores)

merged <- as.matrix(merge(t1,t2, by = 1))
rf <- colorRampPalette(brewer.pal(9,'YlOrRd'))

R2.all.all <- matrix(nrow = 51, ncol = 0)
# Smooth spline
for(j in 3:(nfactors+2)){
  R2.all <- matrix(nrow = 0, ncol = 1)
  ss.all <- list()
  col.all <- matrix(nrow = 0, ncol = 1)
  
  x <- as.numeric(merged[,j]) # factor score
  y <- as.numeric(merged[,2]) # MH score
  
  layout(matrix(1:2,ncol=2), width = c(5,1),height = c(1,1))
  par(bg = "white", col.axis = "black", col.lab = "black", col.main = "black", 
      col.sub = "black", fg = "black", mar = c(3.5,4,1,1), ps = 10, tcl = 0.1, mgp=c(1,0,0))
  plot(x, y, 
       xlab = factorName(colnames(merged)[j]),
       ylab = "Mental health not good for ≥ 14\ndays among adults aged ≥ 18 years", 
       col = alpha("grey50", 0.2), pch = 16, cex = 0.5)
  
  for(k in 1:length(stateCodes)){
    this.State <- merged[which(grepl(stateCodes[k], substr(merged[,1],1,2))),]
    
    this.x <- as.numeric(this.State[,j]) # factor score
    this.y <- as.numeric(this.State[,2]) # MH score
    
    ss <- smooth.spline(this.x, this.y, df = 3, keep.data = T)
    
    TSS <- sum((this.y - mean(this.y))^2)
    RSS <- sum((this.y - predict(ss, this.x)$y)^2)
    R2 <- 1-(RSS/TSS)
    
    R2.all <- rbind(R2.all, R2)
    col.all <- rbind(col.all, rf(100)[100 - (round(R2,2)*100)])
    ss.all[[length(ss.all)+1]] <- ss
  }
  
  
  for(k in order(R2.all, decreasing = F)){
    lines(ss.all[[k]], type = "l", col = col.all[k], lwd = 2.5)
  }
  
  legend_image <- as.raster(matrix(rf(20), ncol=1))
  par(bg = "white", col.axis = "black", col.lab = "black", col.main = "black",
      col.sub = "black", fg = "black", mar = c(0,0,1,0), ps = 10, tcl = 0.1, mgp=c(1,0,0))
  plot(c(0,2),c(0,1),type = 'n', axes = F, xlab = '', ylab = '')
  text(x=0.6, y = seq(.75,0.25,l=11), labels = seq(1,0,l=11), adj = c(0, 0.5))
  text(x = 0, y = 0.8, labels = expression(R^2), font = 2, adj = c(0, 0.5))
  rasterImage(legend_image, 0, 0.25, 0.5,0.75)
  rect(0, 0.25, 0.5,0.75, lwd = 2)
  segments(0, seq(.75,0.25,l=11), x1 =0.55, lwd = 2)
  
  R2.all.all <- cbind(R2.all.all, R2.all)
  colnames(R2.all.all)[j-2] <- factorName(colnames(merged)[j])
}
```
```{r, fig.height=10, fig.width=10}
par(mar = c(4,4,1,1), ps = 10, mgp = c(2,1,0), yaxs = "i")
plot.new()
plot.window(ylim = c(0, 1), xlim = c(0.5,5.5))
rasterImage(legend_image, 0, 0, 6,1)
rect(0, 0, 0.6, 1, col = "white", border = "white")
rect(1.4, 0,1.6, 1, col = "white", border = "white")
rect(2.4, 0,2.6, 1, col = "white", border = "white")
rect(3.4, 0,3.6, 1, col = "white", border = "white")
rect(4.4, 0,4.6, 1, col = "white", border = "white")
rect(5.4, 0,6, 1, col = "white", border = "white")

rect(0.6, 0, 1.4, quantile(R2.all.all[,1])[2], col = "white", border = "white")
rect(1.6, 0, 2.4, quantile(R2.all.all[,2])[2], col = "white", border = "white")
rect(2.6, 0, 3.4, quantile(R2.all.all[,3])[2], col = "white", border = "white")
rect(3.6, 0, 4.4, quantile(R2.all.all[,4])[2], col = "white", border = "white")
rect(4.6, 0, 5.4, quantile(R2.all.all[,5])[2], col = "white", border = "white")

rect(0.6, quantile(R2.all.all[,1])[4], 1.4, 1, col = "white", border = "white")
rect(1.6, quantile(R2.all.all[,2])[4], 2.4, 1, col = "white", border = "white")
rect(2.6, quantile(R2.all.all[,3])[4], 3.4, 1, col = "white", border = "white")
rect(3.6, quantile(R2.all.all[,4])[4], 4.4, 1, col = "white", border = "white")
rect(4.6, quantile(R2.all.all[,5])[4], 5.4, 1, col = "white", border = "white")

boxplot(R2.all.all, ylab = "R^2 value", las = 2, ylim = c(0,1), names = NA, add = T, col = "transparent", border = "black")
labels <- colnames(R2.all.all)
text(c(1:5), -0.02, labels = labels, srt = 15, 
     adj = c(1,1), xpd = TRUE, cex=.9)
```
```{r}
for (i in 1:5){
  print(colnames(R2.all.all)[i])
  print(paste("     First Quartile:", quantile(R2.all.all[,i])[2]))
  print(paste("     Median:", median(R2.all.all[,i])))
  print(paste("     Third Quartile:", quantile(R2.all.all[,i])[4]))
  }
```

  
```{r, fig.height=4, fig.width=4}
par(mar = c(4,4,1,1), ps = 10, mgp = c(2,1,0), yaxs = "i")
plot.new()
plot.window(ylim = c(0, 1), xlim = c(0.5,5.5))
rasterImage(legend_image, 0, 0, 6,1)
rect(0, 0, 0.6, 1, col = "white", border = "white")
rect(1.4, 0,1.6, 1, col = "white", border = "white")
rect(2.4, 0,2.6, 1, col = "white", border = "white")
rect(3.4, 0,3.6, 1, col = "white", border = "white")
rect(4.4, 0,4.6, 1, col = "white", border = "white")
rect(5.4, 0,6, 1, col = "white", border = "white")

rect(0.6, 0, 1.4, quantile(R2.all.all[,1])[2], col = "white", border = "white")
rect(1.6, 0, 2.4, quantile(R2.all.all[,2])[2], col = "white", border = "white")
rect(2.6, 0, 3.4, quantile(R2.all.all[,3])[2], col = "white", border = "white")
rect(3.6, 0, 4.4, quantile(R2.all.all[,4])[2], col = "white", border = "white")
rect(4.6, 0, 5.4, quantile(R2.all.all[,5])[2], col = "white", border = "white")

rect(0.6, quantile(R2.all.all[,1])[4], 1.4, 1, col = "white", border = "white")
rect(1.6, quantile(R2.all.all[,2])[4], 2.4, 1, col = "white", border = "white")
rect(2.6, quantile(R2.all.all[,3])[4], 3.4, 1, col = "white", border = "white")
rect(3.6, quantile(R2.all.all[,4])[4], 4.4, 1, col = "white", border = "white")
rect(4.6, quantile(R2.all.all[,5])[4], 5.4, 1, col = "white", border = "white")

boxplot(R2.all.all, ylab = "R^2 value", las = 2, ylim = c(0,1), names = NA, add = T, col = "transparent", border = "black")
labels <- colnames(R2.all.all)
text(c(1:5), -0.04, labels = labels, srt = 20, 
     adj = c(1,1), xpd = TRUE, cex=.9)
```
  
  
```{r, include=F, eval=F}
for (i in 1:5){
  hist(FA_Results$loadings[,i], breaks = 15, main = factorName(colnames(FA_Results$loadings)[i]))
}

```